Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy

### 2023-10-30 manuscript/manuscript.qmd
Mega updates to results section

---
title: "Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy"
short-title: "Why Donors Donate"
date: "2023-09-09"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
  orcid: 0000-0003-3506-7590
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  Prior research on nonprofit fundraising has primarily examined how financial transparency and organizational accountability influence individual donor behavior. Over the past two decades, however, an increasing number of host governments have cracked down on non-governmental organizations (NGOs). QUESTION. We test these questions by administering a series of conjoint experiments in the US that examine...We find.... This suggests...
keywords:
  - philanthropy
  - civil society
  - conjoint experiments
  - donor heuristics
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/silent-skywalk>"
bibliography: references.bib
link-citations: true
---

@@

```{r libraries-data}
library(tidyverse)
library(targets)

# Targets stuff
tar_config_set(store = here::here('_targets'),
               script = here::here('_targets.R'))

invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
```

Over the last three decades, governments globally have increasingly sought to crack down on NGOs (CITE). Repressive governments have repressed these organizations using both violence, as well as administrative crackdown, which create barriers to funding and advocacy for a variety of NGOs. (CITE) Policymakers have referred to this phenomenon as the "closing of civic space" or "shrinking civic space" (CITE) and data from the CIVICUS Civil Society Monitor shows that civil society groups face serious restrictions in nearly 60% of countries—111 in total (UPDATE FROM NEW CIVICUS DATA). These attacks—most of which have been implemented through anti-NGO laws—have adversely affected the ability of both domestic and international NGOs to operate and raise funds. Official aid funds have decreased substantially in countries that have imposed restrictive anti-NGO legislation [@ChaudhryHeiss:2018; @DupuyPrakash:2018] and official aid donors channel fewer funds to contentious issues and instead choose to fund "tame" causes [@ChaudhryHeiss:2018; @Bush:2015; @HERROLD; @SPRINGMAN]. In response to this dramatic shift in the NGO funding landscape, international NGOs (INGOs) have increasingly turned to foreign individual donors to offset the loss of funding from government and foundation sources [@BanksHulmeEdwards:2015].

The amount of private foreign aid is not insignificant—from 2000–2010, private giving by U.S. entities towards international causes doubled in real terms [@DesaiKharas:2018, 505]. In 2016, U.S. private philanthropy towards international NGOs amounted to \$22.06 billion [@GivingUSA:2017]. Most notably, a large portion of this growth was driven by individuals making small-scale donations. In 2017, individuals gave US\$286.7 billion or approximately 70% of total giving to international affairs. This further increased to US\$427.71 billion in 2018, 85.7% of which was through individuals [@GivingUSA:2019]. This turn to private donors opens INGOs up to new dynamics in seeking out aid. Many organizations that had previously relied on governments and foundations had optimized their fundraising strategies for those donors, but now need to tailor their fundraising strategies to private individuals, who behave differently from governments and foundations.

Individuals have different preferences and rely on signals of effectiveness that differ from the benchmarks that large donor agencies rely on, including financial transparency and accountability. While INGOs have control over these organizational-level factors that motivate donations, the political environment in the host countries INGOs work in—including the global crackdown on civil society that is causing many organizations to seek out private funding in the first place—is a structural factor that INGOs can rarely influence on their own. How do organizational elements of private donor decision-making interact with broader structural limits on INGO fundraising? How do individual donors in the U.S. feel about donating to legally besieged NGOs abroad? How do organizational-level factors such as financial transparency and accountability compare with more structural-level factors such as host country politics when deciding to donate?

@@

These findings make two main contributions. While recent research on private foreign aid has examined funding decisions of donor NGOs and foundations[^1], we know less about the funding choices of the individuals responsible for much of these private aid flows [@DesaiKharas:2018]. The systematic study of private foreign aid is still in its infancy—there is an absence of expansive and high quality data [@GreenhillPrizzonRogerson:2013], and most research has been restricted to elite, high net-worth donors [@USTrust2014]. Our research makes a significant contribution to research on individual giving and indicates the promise and limits of different framing appeals on individual donor preferences.

@@

Substantial research on NGOs has looked at the motivation of individual donors, including altruism, reputational benefits, and alignment with personal values [@BilodeauSlivinski:1997; @Wiepking:2010; @BekkersWiepking:2011]. However, this research has overwhelmingly looked at individual donor behavior and motivation for giving to local—rather than international—causes (CITE -2). Understanding the latter is important as philanthropy to international causes can be harder to motivate because the number of recipients is larger and further removed from the donor [@CasaleBaumann:2015, 100]. Further, as @MicklewrightSchnepf:2009 point out, motivations for donating to ongoing international causes differ from donating money to natural disasters or emergency international causes. Under ordinary circumstances, “donors are more likely to donate to a charity operating locally than to a charity providing identical service abroad” [@Tremblay-BoirePrakash:2017, 644]. However, we lack adequate data and theory regarding philanthropy towards INGOs.

@@

Donors respond to a host of heuristics and signals when deciding whether to donate to an NGO, since seeking out complete information about an organization's deservingness and efficiency is costly and time-consuming [@CrosonShang:2011; @Tremblay-BoirePrakash:2017]. Nonprofits fundraising internationally may therefore rely on framing to control perceptions. Below, we theorize the effect of different frames surrounding a country’s legal environment and nonprofit organizational characteristics on individual donor preferences.

@@

In theory, donors should scrutinize each aspect of an NGO’s organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Instead, individuals rely on heuristics to simplify their decision-making.

@@

In addition to organizational factors discussed above, political contexts in NGO host countries, especially their environments for civil society organizations, can also influence donor preferences for donating to foreign NGOs. A global cascade of anti-NGO laws in recent decades has created barriers to entry, funding, and advocacy for civic organizations [@GlasiusSchalkDeLange:2020; @Schmidt-et-al; @Chaudhry:2022; @Heiss:2017; @DupuyRonPrakash:2015]. As a result, NGOs must spend more time, effort, and resources to ensure their survival -- at the expense of pursuing their missions. Some NGOs have adapted to their repressive environments by recreating their organizations and changing their mission to avoid directly confronting the government. Analyzing the impact of anti-NGO laws in Bangladesh and Zambia, scholars have similarly found that NGOs with a broad mission ended up restricting themselves to service work and giving up on their advocacy work. Those that continued to engage in transnational advocacy changed the targets of their advocacy, the issues they focused on, and the language they used. For instance, many NGOs stopped lobbying the central government and focused their advocacy efforts on lower levels of government [@FransenDupuyHinfelaar:2021]. Similarly, in Russia, many NGOs working with foreign partners and lobbying the central government no longer consider it an effective strategy [@SundstromHenrySperling:2022]. Overall, fewer groups may end up focusing on contentious causes, instead focusing on regime-compatible causes.

The changes in civic environments impacting NGO activity have also influenced official donor responses. In response to the passage of these laws, scholars have found that while multilateral donors do not reduce aid following the passage of anti-NGO laws, they are associated with a 32% decline in bilateral aid inflows [@DupuyPrakash:2018]. Further, advocacy-oriented donors -- i.e., those that fund democracy and civil society promotion activities -- reduced their spending by 74%, but did not cut their spending on development projects such as education, health, water, and sanitation [@RightSpringmanWibbels:2022]. However, previous literature has established that individual private donors may not necessarily use the same metrics as official donors (CITE). Private foreign aid does not have to deal with the same strategic and political considerations as official aid and therefore, may be in a better position to respond to recipients' needs on the grounds [@EasterlyWilliamson:2011]. These anti-NGO laws may signal to private donors that NGOs undertake crucial work abroad, which is why governments perceive them as threatening and crack down on them. Therefore, we expect that government restrictions abroad will increase respondents' likelihood of donating to the organization.

>**H~2~**: If NGOs face legal crackdowns abroad, then individual donors will be more likely to donate to them.

@@

```
![Causal diagram showing causes of donor behavior. D = donation to NGO (outcome); G = NGO-government relationship, I = NGO issue area, F = NGO funding source, O = NGO organizational and managerial practices.](output/figures/causal-path.pdf){#fig:big-dag}
```

Identifying the causal link between structural factors and donor behavior is more difficult than measuring the link between organizational factors and donor behavior, however, because (1) NGOs have less direct control over their political environments, and (2) organizational characteristics like financial transparency and accountability can sour the NGO-government relationship -- governments may be more likely to target NGOs that lack financial transparency and accountability. @fig:big-dag helps disentangle this complex relationship. Organizational factors (**O**, or organizational practices like financial disclosure and accountability practices; **I**, or issue area; and **F**, or funding) all influence the decision to donate (**D**). Structural factors like an NGO's relationship with its host government (**G**) also influence donation behavior, but the relationship itself is shaped by both transparency and accountability. As such, legal crackdowns do not occur in isolation. Countries are more likely to repress NGOs when organizational features are in tension with government preferences and when organizations receive substantial funding from abroad [@DupuyRonPrakash:2015]. Therefore, the same organizational characteristic that can increase the likelihood of crackdown can also simultaneously serve as a donor heuristic and influence perceptions of NGO deservingness.

@@

Since we are running a conjoint experiment, the interactions between the main treatment variables described above is simply an exploratory analysis. Importantly, when testing these hypotheses, we do not assume that individual donors have perfect knowledge about an NGO or the country an NGO operates in. Moreover, the average donor will not know exactly what “crackdown” entails for any given organization (i.e., does it mean an NGO was expelled, or that its assets were frozen, or that it received a fine?). For the sake of this experiment, this ambiguity is not a central concern as the heuristics we explore here are explicitly simplified short- cuts to donor information-gathering. Knowing that an organization is financially transparent (even if donors are not familiar with specific kinds of requirements regarding transparency) or knowing that an NGO faces legal trouble abroad should be enough to trigger the effect of these frames on donor preferences. Moreover, although many NGOs make information about their operations accessible through annual reports, it is time-consuming to track down that information. In each of these hypotheses, we assume that donors will not spend additional effort to research the exact details concerning crackdown, transparency and accountability. Instead, donors rely on signals and heuristics about the organization to shape their preferences and behavior [@SzperPrakash:2011].

# Methods

## Conjoint analysis

It is possible to study the effects of both organizational and structural heuristics on donor decision making individually, but disentangling the interaction of the two heuristics from individual donor characteristics adds complex dimensionality and makes more standard experimental work difficult, costly, and statistically fraught. To address this, we use conjoint analysis to simultaneously test individual donor preferences in the face of both organizational and structural heuristics. Conjoint analysis is a staple in marketing research for evaluating which product attributes are most important to consumers. It has been extensively used in new product development and product line optimization. Conjoint analysis is generally applicable whenever individuals face complex, multi-dimensional tradeoffs that make it difficult (or even impossible) to explicitly elicit preferences. In the context of this project, we use conjoint techniques to evaluate donor rather than consumer preferences as they choose NGOs which to donate to based on both organizational and structural heuristics.

Conjoint analysis has been widely used within social sciences to explain causality and has recently begun to be used in political and public policy fields. @Helfer:2016 and @Mummolo:2016 both used conjoint analysis to assess perceptions of political communications. @KnudsenJohannesson:2019 argue that political communication research requires conjoint analysis because it accounts for the masking effects of factors as well as the complexity of a respondent's judgment of credibility. Because we also want to capture the complexity of a respondent's judgment of which NGO they should donate to, if any, we saw potential to expand the application of conjoint analysis within public policy.

@@

After collecting baseline information on respondent demographics and attitudes toward charity and voluntarism, we presented respondents with four randomly selected international nonprofit organizations with four randomly assigned features.^[See the appendix for a copy of the survey.] We included organizations that are associated with a range of contentious and noncontentious issues: Amnesty International, Greenpeace, Oxfam, and the International Committee of the Red Cross. We chose these 4 organizations as the represent the core sectors of activity according to Charity Navigator (Environment, Human Rights, Humanitarian Relief, and Development). We then displayed various combinations of features, including one of four issue areas (emergency response, environmental advocacy, human rights advocacy, and refugee relief), one of two organizational practices (financial transparency and third-party accountability audits), one of three funding sources (many small private donations, a handful of wealthy private donors, and government grants), and one of three relationships with host governments (friendly, criticized by the government, and under government crackdown). We presented twelve sets of four hypothetical combinations of organizational features, and because our partial fractional factorial design resulted in 288 (4 × 4 × 2 × 3 × 3) possible combinations of organizations and features, no single respondent was offered every combination.

Survey participants were recruited through Centiment, a commercial online provider of high quality non-probability opt-in survey panels. Centiment ensures panel quality by actively recruiting representative samples of the US population and provides monetary incentives and rewards to participants. To see how varying NGO characteristics influence the decision to donate, our sample was representative of the population of people who are likely to donate to charity. We asked potential participants a screening question early in the survey ("Q2.5: How often do you donate to charity?"). If a participant responded that they give once every few years or never, they were disqualified from the study. We received 1,016 viable responses, which is a sufficient size for model estimation. A sample size of at least 500 respondents is typical for estimating a hierarchical Bayesian model based on conjoint data. We doubled this amount because we are interested in analyzing subpopulations of respondents, which requires a larger sample.

```
output/tables/tbl-sample-details.md
```

@tbl:sample-details provides a detailed breakdown of the individual characteristics of our sample. In general, respondents were well balanced across genders and tended to be older than the national median age of 36—less than 20% of respondents were under 36 years old. The majority of respondents (45%) had never been married, while nearly 40% were married, with 10% divorced. A quarter of the sample had a high school degree, with the remainder attending some level of college or university. The majority (58%) earned less than the national median income (\$61,372).

Since we limited the sample to people who frequently donate to charities, most respondents have a favorable opinion of nonprofits and charities, with more than half reporting a 5 or greater (on a 7-point scale) of trust in charities. Nearly 90% of respondents indicated that it is important to trust charities, rating importance of trust at least a 3 on a 7-point scale. The majority of respondents volunteer at least once a year, donate to charity at least once a year, and give between \$1–\$999 annually. Most rarely attend church, but feel that religion is important. Respondents are also engaged civically (three quarters voted in the last election) and follow national news closely (70% follow the news daily). Respondents are fairly balanced ideologically, with the majority reporting a 3, 4, or 5 on a 7-point scale of liberal–conservative political ideology.

## Modeling

@@

>>>>

---
title: "Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy"
short-title: "Why Donors Donate"
date: "2023-10-29"
author:
- name: Suparna Chaudhry
  affiliation: Lewis and Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
  orcid: 0000-0003-3506-7590
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
  orcid: 0000-0002-3948-3914
abstract: |
  Prior research on nonprofit fundraising has primarily examined how financial transparency and organizational accountability influence individual donor behavior. Over the past two decades, however, an increasing number of host governments have cracked down on non-governmental organizations (NGOs). QUESTION. We test these questions by administering a series of conjoint experiments in the US that examine...We find.... This suggests...
keywords:
  - philanthropy
  - civil society
  - conjoint experiments
  - donor heuristics
code-repo: "Access the code, data, and analysis at <https://github.com/andrewheiss/silent-skywalk>"
bibliography: references.bib
link-citations: true
---

@@

```{r libraries-data}
library(tidyverse)
library(targets)
library(tidybayes)
library(patchwork)
library(ggforce)
library(scales)
library(glue)
library(gt)
library(gtExtras)
library(here)

# Targets stuff
tar_config_set(
  store = here("_targets"),
  script = here("_targets.R")
)

tar_load(c(grid_treatment_only, level_lookup, feature_lookup))
preds_all <- tar_read(preds_conditional_treatment_only)

invisible(list2env(tar_read(graphic_functions), .GlobalEnv))
invisible(list2env(tar_read(table_functions), .GlobalEnv))

theme_set(theme_ngo())
```

Over the last three decades, governments globally have increasingly sought to crack down on NGOs (CITE). Repressive governments have repressed these organizations using both violence, as well as administrative crackdown, which create barriers to funding and advocacy for a variety of NGOs. (CITE) Policymakers have referred to this phenomenon as the "closing of civic space" or "shrinking civic space" (CITE) and data from the CIVICUS Civil Society Monitor shows that civil society groups face serious restrictions in nearly 60% of countries---111 in total (UPDATE FROM NEW CIVICUS DATA). These attacks---most of which have been implemented through anti-NGO laws---have adversely affected the ability of both domestic and international NGOs to operate and raise funds. Official aid funds have decreased substantially in countries that have imposed restrictive anti-NGO legislation [@ChaudhryHeiss:2018; @DupuyPrakash:2018] and official aid donors channel fewer funds to contentious issues and instead choose to fund "tame" causes [@ChaudhryHeiss:2018; @Bush:2015; @Herrold:2020; @RightSpringmanWibbels:2022]. In response to this dramatic shift in the NGO funding landscape, international NGOs (INGOs) have increasingly turned to foreign individual donors to offset the loss of funding from government and foundation sources [@BanksHulmeEdwards:2015].

The amount of private foreign aid is not insignificant---from 2000--2010, private giving by U.S. entities towards international causes doubled in real terms [@DesaiKharas:2018, 505]. In 2016, U.S. private philanthropy towards international NGOs amounted to \$22.06 billion [@GivingUSA:2017]. Most notably, a large portion of this growth was driven by individuals making small-scale donations. In 2017, individuals gave US\$286.7 billion or approximately 70% of total giving to international affairs. This further increased to US\$427.71 billion in 2018, 85.7% of which was through individuals [@GivingUSA:2019]. This turn to private donors opens INGOs up to new dynamics in seeking out aid. Many organizations that had previously relied on governments and foundations had optimized their fundraising strategies for those donors, but now need to tailor their fundraising strategies to private individuals, who behave differently from governments and foundations.

Individuals have different preferences and rely on signals of effectiveness that differ from the benchmarks that large donor agencies rely on, including financial transparency and accountability. While INGOs have control over these organizational-level factors that motivate donations, the political environment in the host countries INGOs work in---including the global crackdown on civil society that is causing many organizations to seek out private funding in the first place---is a structural factor that INGOs can rarely influence on their own. How do organizational elements of private donor decision-making interact with broader structural limits on INGO fundraising? How do individual donors in the U.S. feel about donating to legally besieged NGOs abroad? How do organizational-level factors such as financial transparency and accountability compare with more structural-level factors such as host country politics when deciding to donate?

@@

These findings make two main contributions. While recent research on private foreign aid has examined funding decisions of donor NGOs and foundations[^1], we know less about the funding choices of the individuals responsible for much of these private aid flows [@DesaiKharas:2018]. The systematic study of private foreign aid is still in its infancy---there is an absence of expansive and high quality data [@GreenhillPrizzonRogerson:2013], and most research has been restricted to elite, high net-worth donors [@USTrust2014]. Our research makes a significant contribution to research on individual giving and indicates the promise and limits of different framing appeals on individual donor preferences.

@@

Substantial research on NGOs has looked at the motivation of individual donors, including altruism, reputational benefits, and alignment with personal values [@BilodeauSlivinski:1997; @Wiepking:2010; @BekkersWiepking:2011]. However, this research has overwhelmingly looked at individual donor behavior and motivation for giving to local---rather than international---causes (CITE -2). Understanding the latter is important as philanthropy to international causes can be harder to motivate because the number of recipients is larger and further removed from the donor [@CasaleBaumann:2015, 100]. Further, as @MicklewrightSchnepf:2009 point out, motivations for donating to ongoing international causes differ from donating money to natural disasters or emergency international causes. Under ordinary circumstances, "donors are more likely to donate to a charity operating locally than to a charity providing identical service abroad" [@Tremblay-BoirePrakash:2017, 644]. However, we lack adequate data and theory regarding philanthropy towards INGOs.

@@

Donors respond to a host of heuristics and signals when deciding whether to donate to an NGO, since seeking out complete information about an organization's deservingness and efficiency is costly and time-consuming [@CrosonShang:2011; @Tremblay-BoirePrakash:2017]. Nonprofits fundraising internationally may therefore rely on framing to control perceptions. Below, we theorize the effect of different frames surrounding a country's legal environment and nonprofit organizational characteristics on individual donor preferences.

TODO: Cite @ChaudhryHeiss:2021 somewhere

@@

In theory, donors should scrutinize each aspect of an NGO's organizational structure and programmatic performance prior to donation, but this rarely happens in practice [@Tremblay-BoirePrakash:2017]. Instead, individuals rely on heuristics to simplify their decision-making.

@@

In addition to organizational factors discussed above, political contexts in NGO host countries, especially their environments for civil society organizations, can also influence donor preferences for donating to foreign NGOs. A global cascade of anti-NGO laws in recent decades has created barriers to entry, funding, and advocacy for civic organizations [@GlasiusSchalkDeLange:2020; @BakkeMitchellSmidt:2020; @Chaudhry:2022; @Heiss:2017; @DupuyRonPrakash:2015]. As a result, NGOs must spend more time, effort, and resources to ensure their survival—at the expense of pursuing their missions. Some NGOs have adapted to their repressive environments by recreating their organizations and changing their mission to avoid directly confronting the government. Analyzing the impact of anti-NGO laws in Bangladesh and Zambia, scholars have similarly found that NGOs with a broad mission ended up restricting themselves to service work and giving up on their advocacy work. Those that continued to engage in transnational advocacy changed the targets of their advocacy, the issues they focused on, and the language they used. For instance, many NGOs stopped lobbying the central government and focused their advocacy efforts on lower levels of government [@FransenDupuyHinfelaar:2021]. Similarly, in Russia, many NGOs working with foreign partners and lobbying the central government no longer consider it an effective strategy [@SundstromHenrySperling:2022]. Overall, fewer groups may end up focusing on contentious causes, instead focusing on regime-compatible causes.

The changes in civic environments impacting NGO activity have also influenced official donor responses. In response to the passage of these laws, scholars have found that while multilateral donors do not reduce aid following the passage of anti-NGO laws, they are associated with a 32% decline in bilateral aid inflows [@DupuyPrakash:2018]. Further, advocacy-oriented donors---i.e., those that fund democracy and civil society promotion activities---reduced their spending by 74%, but did not cut their spending on development projects such as education, health, water, and sanitation [@RightSpringmanWibbels:2022]. However, previous literature has established that individual private donors may not necessarily use the same metrics as official donors (CITE). Private foreign aid does not have to deal with the same strategic and political considerations as official aid and therefore, may be in a better position to respond to recipients' needs on the grounds [@EasterlyWilliamson:2011]. These anti-NGO laws may signal to private donors that NGOs undertake crucial work abroad, which is why governments perceive them as threatening and crack down on them. Therefore, we expect that government restrictions abroad will increase respondents' likelihood of donating to the organization.

> **H~2~**: If NGOs face legal crackdowns abroad, then individual donors will be more likely to donate to them.

@@

Identifying the causal link between structural factors and donor behavior is more difficult than measuring the link between organizational factors and donor behavior, however, because (1) NGOs have less direct control over their political environments, and (2) organizational characteristics like financial transparency and accountability can sour the NGO-government relationship---governments may be more likely to target NGOs that lack financial transparency and accountability. Legal restrictions do not occur in isolation. Countries are more likely to repress NGOs when organizational features are in tension with government preferences and when organizations receive substantial funding from abroad [@DupuyRonPrakash:2015]. Therefore, the same organizational characteristic that can increase the likelihood of crackdown can also simultaneously serve as a donor heuristic and influence perceptions of NGO deservingness.

@@

As a result, we are uncertain about the interaction between transparency, accountability, and host country legal environment. We suspect that there is a relationship, but we are uncertain about the exact direction or magnitude. Accordingly, we do not propose any confirmatory hypothesis testing, but instead explore the interaction between our primary treatment variables to inform future research on the interplay between organizational and structural dynamics.

Importantly, when testing these hypotheses, we do not assume that individual donors have perfect knowledge about an NGO or the country an NGO operates in. Moreover, the average donor will not know exactly what "crackdown" entails for any given organization (i.e., does it mean an NGO was expelled, or that its assets were frozen, or that it received a fine?). For the sake of this experiment, this ambiguity is not a central concern as the heuristics we explore here are explicitly simplified short- cuts to donor information-gathering. Knowing that an organization is financially transparent (even if donors are not familiar with specific kinds of requirements regarding transparency) or knowing that an NGO faces legal trouble abroad should be enough to trigger the effect of these frames on donor preferences. Moreover, although many NGOs make information about their operations accessible through annual reports, it is time-consuming to track down that information. In each of these hypotheses, we assume that donors will not spend additional effort to research the exact details concerning crackdown, transparency and accountability. Instead, donors rely on signals and heuristics about the organization to shape their preferences and behavior [@SzperPrakash:2011].

# Research design

It is possible to study the effects of both organizational and structural heuristics on donor decision making individually, but disentangling the interaction of the two heuristics from individual donor characteristics adds complex dimensionality and makes more standard experimental work difficult, costly, and statistically fraught. To address this, we use conjoint analysis to simultaneously test individual donor preferences in the face of both organizational and structural heuristics. Conjoint analysis is generally applicable whenever individuals face complex, multi-dimensional tradeoffs that make it difficult (or even impossible) to explicitly elicit preferences. Conjoint designs have been used extensively in political science to study complex preferences for elections, laws, public policy, and public opinion [@HainmuellerHopkinsYamamoto:2014; @Mummolo:2016; @KertzerRenshonYarhi-Milo:2021; @BansakHainmuellerHopkins:2021]. @KnudsenJohannesson:2019 argue that research related to respondent opinions requires conjoint analysis because it accounts for the masking effects of factors as well as the complexity of a respondent's judgment of credibility. Because we also want to capture the complexity of a respondent's judgment of philanthropic preferences, we use a conjoint design to evaluate the effect of organizational and structural heuristics on philanthropic preferences.

A conjoint research design also allows us to calculate treatment effects even if respondents do not see every combination of organization attributes, and correspondingly provides a substantial increase in statistical power. In standard factorial experimental designs, every combination of treatment attributes would need to be shown to some number of participants would ordinarily need to be shown all possible combinations of experimental attributes. In our experiment, we presented respondents with six possible treatments with varying counts of attributes for each treatment, yielding 576 possible unique combinations of features. Even with respondents answering twelve iterations of the experiment, not every combination was seen. However, as long as all the possible organizational attributes are well randomized and there are no systematic biases toward specific options (i.e. more respondents select the first option because it is the first) or toward earlier iterations of attribute choices (i.e. respondents are more careful and attentive for the first hypothetical organization than the last), we can pool all observations together for specific attributes of interest while marginalizing across all other attributes [@KertzerRenshonYarhi-Milo:2021]. This allows us to (1) estimate the effect of each experimental treatment even if some unique combinations of attributes were unseen, and (2) use a much smaller sample size than would be required in a more traditional factorial design.

While the ability to select key attributes and marginalize over others provides us with analytic flexibility, it also raises possible issues with multiple comparisons, *p*-hacking, and selective cherry-picking [@BansakHainmuellerHopkins:2021]. As such, before launching the survey experiment, we preregistered a set of confirmatory and exploratory hypotheses at the Open Science Framework. Our preregistration protocol is available at `ANONYMIZED_URL` and in Appendix TODO, and our data and reproducible code is available at `ANONYMIZED_URL`.

@@

We recruited survey participants through Centiment, a commercial online provider of high quality non-probability opt-in survey panels. Centiment ensures panel quality by actively recruiting representative samples of the US population and provides monetary incentives and rewards to participants. To see how varying NGO characteristics influence the decision to donate, our sample was representative of the population of people who are likely to donate to charity. We asked potential participants a screening question early in the survey ("Q2.5: How often do you donate to charity?"). If a participant responded that they give once every few years or never, they were disqualified from the study. We also included an attention check question early in the survey and removed respondents who failed the question. Importantly, these screening questions were presented *prior to* the experimental manipulation to avoid post-treatment bias [@MontgomeryNyhanTorres:2018].

Following screening, we received 1,016 viable responses. @tbl:sample-details provides a detailed breakdown of the individual characteristics of our sample. In general, respondents were well balanced across genders and tended to be older than the national median age of 36---less than 20% of respondents were under 36 years old. The majority of respondents (45%) had never been married, while nearly 40% were married, with 10% divorced. A quarter of the sample had a high school degree, with the remainder attending some level of college or university. The majority (58%) earned less than the national median income (\$61,372). Since we limited the sample to people who frequently donate to charities, most respondents have a favorable opinion of nonprofits and charities, with more than half reporting a 5 or greater (on a 7-point scale) of trust in charities. Nearly 90% of respondents indicated that it is important to trust charities, rating importance of trust at least a 3 on a 7-point scale. The majority of respondents volunteer at least once a year, donate to charity at least once a year, and give between \$1--\$999 annually. Most rarely attend church, but feel that religion is important. Respondents are also engaged civically (three quarters voted in the last election) and follow national news closely (70% follow the news daily). Respondents are fairly balanced ideologically, with the majority reporting a 3, 4, or 5 on a 7-point scale of liberal--conservative political ideology.

## Experimental design

After collecting baseline information on respondent demographics and attitudes toward charity and voluntarism, we presented respondents with three randomly generated hypothetical international nonprofit organizations with six randomly assigned features. In this paper, we hypothesize and empirically test the effect of financial transparency, accountability, and host country relationship on the likelihood of donations. In our experiment, we included a few additional treatments to aid with the interpretation and comparison of effect sizes for our treatments of interest. In total, we varied six different organizational and structural attributes that might have an effect on donor behavior: (1) organization name and branding, (2) organization issue area, (3) financial transparency practices, (4) accountability practices, (5) funding sources, and (6) relationship with host government (see @tbl-design).

```{css, echo=FALSE}
/* Remove padding on lists in tables */
.gt_table ul, .gt_table ol {
  padding-left: 1rem;
}
```

```{r tbl-design, include=TRUE}
#| tbl-cap: "All experimental features and levels included in the conjoint experiment"
transparency <- c("Doesn't engage in transparency", "Engages in transparency")
accountability <- c("Doesn't engage in accountability", "Engages in accountability")
govt <- c("Friendly relationship with government", "Criticized by government", "Under government crackdown")
orgs <- c("Amnesty International", "Greenpeace", "Oxfam", "Red Cross")
issues <- c("Emergency response", "Environment", "Human rights", "Refugee relief")
funding <- c("Funded primarily by many small private donations", "Funded primarily by a handful of wealthy private donors", "Funded primarily by government grants")

tribble(
  ~type, ~feature, ~levels,
  "H~1a~", "Financial transparency", paste0(paste0("- ", transparency), collapse = "\n"),
  "H~1b~", "Accountability", paste0(paste0("- ", accountability), collapse = "\n"),
  "H~2~", "Relationship with host government", paste0(paste0("- ", govt), collapse = "\n"),
  "Comparison", "Organization", paste0(paste0("- ", orgs), collapse = "\n"),
  "Comparison", "Issue area", paste0(paste0("- ", issues), collapse = "\n"),
  "Comparison", "Funding source", paste0(paste0("- ", funding), collapse = "\n")
) %>% 
  gt() %>% 
  cols_label(feature = "Feature/Attribute", levels = "Levels", type = "Purpose") %>% 
  fmt_markdown(columns = c(type, levels)) %>% 
  tab_style(
    style = cell_text(v_align = "top"),
    locations = cells_body()
  ) %>% 
  opts_theme()
```

Charity Navigator—a large database of U.S. nonprofit financial information—categorizes its nonprofit organizations into four core sectors: humanitarian relief, human rights, development, and the environment. Accordingly, we used four nonprofits that are stereotypical for each issue: the International Committee of the Red Cross (humanitarian relief), Amnesty International (human rights), Oxfam (development), and Greenpeace (environment). We also varied several other features, including four issue areas (emergency response, environmental advocacy, human rights advocacy, and refugee relief), two organizational practices (financial transparency and third-party accountability audits), three funding sources (many small private donations, a handful of wealthy private donors, and government grants), and three relationships with host governments (friendly, criticized by the government, and under government crackdown).

After presenting respondents with three organizations, we asked them which of the three they would be willing to donate to, along with an option for no selection (see @tbl-example for an example). We then repeated the process eleven more times for a total of twelve randomized iterations of hypothetical combinations of attributes, resulting in 12,192 completed experimental tasks (12 iterations × 1,016 respondents).

```{r tbl-example, include=TRUE}
#| tbl-cap: "Example possible experimental task, with three randomized combinations of organizational attributes"
withr::with_seed(12345, {
  tribble(
    ~Attribute, ~`Option 1`, ~`Option 2`, ~`Option 3`, ~None,
    "Organization", sample(orgs, 1), sample(orgs, 1), sample(orgs, 1), "—",
    "Issue area", sample(issues, 1), sample(issues, 1), sample(issues, 1), "—",
    "Transparency", sample(transparency, 1), sample(transparency, 1), sample(transparency, 1), "—",
    "Accountability", sample(accountability, 1), sample(accountability, 1), sample(accountability, 1), "—",
    "Funding sources", sample(funding, 1), sample(funding, 1), sample(funding, 1), "—",
    "Relationship with host government", sample(govt, 1), sample(govt, 1), sample(govt, 1), "—"
  ) %>%
    gt() %>%
    cols_align(align = "center", columns = everything()) %>%
    cols_align(align = "left", columns = Attribute) %>%
    cols_width(everything() ~ pct(20)) %>%
    tab_style(
      style = cell_text(v_align = "top"),
      locations = cells_body()
    ) %>%
    tab_style(
      style = cell_text(weight = "bold"),
      location = cells_body(columns = Attribute)
    ) %>%
    opts_theme()
})
```

## Modeling and estimands

We analyze the results using a multilevel Bayesian multinomial model. Our experimental data has a natural hierarchical structure, with 3 questions nested inside 12 separate experimental tasks, nested inside each of the 1,016 respondents, which lends itself to multilevel modeling. Relying on standard OLS regression can discard useful information about (1) the relationships and covariance between the different combinations of feature levels offered to respondents, and (2) individual-specific differences in how respondents react to different feature levels [@JensenMarbleScheve:2021]. Since it was impossible for every respondent to see every possible all 12,000 experimental tasks, multilevel modeling allows us to pool together information about respondents with similar characteristics facing similar sets of choices. Moreover, using random respondent effects provides natural regularization and shrinkage for our estimates—experimental tasks that happened to appear more often due to chance will be accounted for and not influence the overall causal effect.

@@

# References
